Resolvable Ambiguity

### 2019-03-26 paper/paper.tex
New figures, lightly edited later sections

This section introduces an alternative (isomorphic) definition of a word that will be used to motivate the correctness of later algorithms. The modifications permitted by the previous section significantly limit how $G'$ can differ from $G$. In particular, no new terminals are introduced, except '(' and ')', and they are always introduced in a well-balanced fashion.

@@

We will now note a few things about the words $w' \in \words(t)$ for some given $t \in L(T)$:

\begin{itemize}
\item All $w'$ have the same basic word. Limiting the modifications to those in the previous section means that they can only differ in parentheses. This also implies that two trees that share a word (i.e., that can be ambiguous) must also share a basic word.
\item Some parentheses are required, i.e., some ranges must be present in all $w'$. For example, removing the parentheses in ''$(1 + 2) * 3$'' changes the tree produced, since multiplication has higher precedence than addition.
\item There is a finite set of possible ranges in the range bags of $w'$. Grouping parentheses can only be added if they exactly cover a node in the parse tree, and other parentheses can only be added where $G$ allows them.
\item Duplicated grouping parentheses do not matter. For example, $((1 + 2)) * 3$ permits the same syntax trees as $(1 + 2) * 3$. If there are no parentheses present in $G$ then all parentheses are grouping parentheses, thus we can consider the range bag as a \emph{set} instead.
\end{itemize}

@@

This section describes a decision algorithm for detecting unresolvable ambiguities in a pair of grammars $G$ and $G'$, starting with a version with several limitations, most of which are later lifted. The overarching goal is to find a tree $t \in L(T)$ such that there is no $w' \in L(G')$ for which $\parse(w') = \{t\}$. Alternatiely, find a tree that has no unambiguous words.

@@

\begin{enumerate}
  \item $G$ contains no parentheses, i.e., given $G = (S, P, \NT, \T, \I)$ we require $\T \cap \{\verb|'('|, \verb|')'|\} = \emptyset$. This implies that all parentheses in $G'$ are grouping parentheses, thus we only need to consider words $w' \in L(G')$ whose range bag is a set.

  \item $G'$ is constructed by adding parentheses to all non-terminals of $G$, with no other modifications, i.e., $G' = (S, P', \NT, \T \cup \{\verb|'('|, \verb|')'|\}, \I \cup \NT)$ where $P' = P \cup \{ N: N -> \verb|'('|N\verb|')'| \mid N \in \NT \}$. This implies that there are no required grouping parentheses.

  \item No production has a right-hand side that matches a single non-terminal, i.e., $\forall (i:N->r) \in P.\ \NT \cap R(r) = \emptyset$\footnote{$R$ is here the more traditional language of a regular expression, i.e., $R(r) \subseteq (\T \cup \NT)^{*}$. It should be written more explicitly somewhere, but I'm lazy at the moment.}.
\end{enumerate}

\noindent The first two limitations allow us to place all words in $\words(t)$ for some given $t$ in a lattice, whose structure is formed by the subset ordering of the range sets. For example, the tree in Figure~\ref{fig:example-grammar:sem-tree} (but where $G'$ has no associativity or precedence) has the lattice of words seen in Figure~\ref{fig:lattice} (omitting the pair of parentheses around the entire word, since they are always permissible).

\begin{figure*}[t]
\begin{tikzpicture}%[scale = 0.7,font=\tiny]
  \node (a) at (0,5) {$((1) + (2)) * (3)$};

  \node (b1) at (-3.6,4.2) {$(1) + (2) * (3)$};
  \node (b2) at (-1.2,4.2) {$((1) + (2)) * 3$};
  \node (b3) at ( 1.2,4.2) {$((1) + 2) * (3)$};
  \node (b4) at ( 3.6,4.2) {$(1 + (2)) * (3)$};

  \node (c1) at (-5.5,3) {$(1) + (2) * 3$};
  \node (c2) at (-3.3,3) {$(1) + 2 * (3)$};
  \node (c3) at (-1.1,3) {$((1) + 2) * 3$};
  \node (c4) at (1.1,3) {$1 + (2) * (3)$};
  \node (c5) at (3.3,3) {$(1 + (2)) * 3$};
  \node (c6) at (5.5,3) {$(1 + 2) * (3)$};

  \node (d1) at (-3.6,1.8) {$(1) + 2 * 3$};
  \node (d2) at (-1.2,1.8) {$1 + (2) * 3$};
  \node (d3) at ( 1.2,1.8) {$1 + 2 * (3)$};
  \node (d4) at ( 3.6,1.8) {$(1 + 2) * 3$};

  \node (e) at (0, 1) {$1 + 2 * 3$};

  \draw (d1) -- (c1);
  \draw (d1) -- (c2);
  \draw (d1) -- (c3);
  \draw (d2) -- (c1);
  \draw (d2) -- (c4);
  \draw (d2) -- (c5);
  \draw (d3) -- (c2);
  \draw (d3) -- (c4);
  \draw (d3) -- (c6);
  \draw (d4) -- (c3);
  \draw (d4) -- (c5);
  \draw (d4) -- (c6);

  \draw (c1) -- (b1);
  \draw (c1) -- (b2);
  \draw (c2) -- (b1);
  \draw (c2) -- (b3);
  \draw (c3) -- (b2);
  \draw (c3) -- (b3);
  \draw (c4) -- (b1);
  \draw (c4) -- (b4);
  \draw (c5) -- (b2);
  \draw (c5) -- (b4);
  \draw (c6) -- (b3);
  \draw (c6) -- (b4);

  \draw (b1) -- (a);
  \draw (b2) -- (a);
  \draw (b3) -- (a);
  \draw (b4) -- (a);
%%   \draw[preaction={draw=white, -,line width=6pt}] (a) -- (e) -- (c);
\end{tikzpicture}
\caption{The lattice of words for the tree in Figure~\ref{fig:example-grammar:sem-tree}, assuming no precedence in $G'$.}
\label{fig:lattice}
\end{figure*}

The bottom word corresponds to the empty range set, while the top word corresponds to the set of possible parentheses ranges. All trees that can be ambiguous must thus share the same bottom word, while the top may differ.

@@

\begin{itemize}
\item The top word is unambiguous, then this is not a tree we are looking for.
\item It is ambiguous. That means that there is some other lattice for some other tree that also contains the word. Call the top word of this other lattice $w'_2$. For $w'$ to be in the lattice whose top is $w'_2$ its rangeset must be a subset of the rangeset of the latter. But this is true also for all the other words in the lattice, thus there are no unambiguous words in the lattice, i.e., this tree is what we are looking for.
\end{itemize}

The algorithm thus looks for two trees, where the set of possible parentheses for one is a superset of the set of possible parentheses for the other. To do this, we use a linear representation of each lattice: namely the top word. If two trees have the same top word, that implies that their sets of possible parentheses are equal. If we can add parentheses to one word and get the other word, that implies that the rangeset of the former is a subset of the latter.
that term later

We will now construct a pushdown automaton that recognizes these top words in such a way that there is a bijection between successful runs and trees in $L(T)$. As a running example, we will use the following (very simple) grammar $G$:

\setlength{\grammarindent}{3.5em}
\begin{grammar}
  <N> ::= succ: 's' <N>
    \alt zero: 'z'
\end{grammar}

@@

\noindent Intuitively, we parse a child node, but put parentheses around it. This is where we use assumption 3, without it we might introduce double parentheses here.

@@

We now have a pushdown automaton (call it $A_{()}$) that recognizes the top word for each tree in $L(T)$. Next we need to be able to add arbitrary parentheses, to produce a word with a rangeset superset. To do this we create a copy of $A_{()}$ with one modification: for every state $s$ in $A_{()}$ (except the initial and final states), add two transitions $s \xrightarrow{'(', +p} s$ and $s \xrightarrow{')', -p} s$. To avoid cluttering the graph too much, these transitions are shown unlabeled below:

@@

We will call this automaton $A'_{()}$. Successful runs in this automaton have a surjection to runs in $A_($ (ignore transitions along the newly added edges), and thus also have a surjection to parse trees in $L(T)$. We can also note that every sucessful run in $A_{()}$ is also a successful run in $A'_{()}$, since the latter has all states and transitions of the former. Furthermore, two distinct runs, one in $A_{()}$ (call it $p$) and one in $A'_{()}$ (call it $p'$), that both recognize the same word must represent different parse trees in $L(T)$. To see why, we consider two cases:

@@

Two distinct successful runs that accept the same word thus represent two trees where one permits a superset rangeset of the other. To find such runs we construct a product automaton and trim it. We can construct a product automaton since both $A_{()}$ and $A'_{()}$ are visibly pushdown, with the same partition of the input alphabet (push on open parenthesis, pop on close parenthesis, do nothing otherwise). We can trim the product since it retains the same partitioning and thus is also visibly pushdown.

@@

>>>>

This section introduces an alternative (isomorphic) definition of a word that will be used to motivate the correctness of later algorithms. The method by which we generate $G_w$ and $G'_w$ limits the possible differences between them significantly. In particular, no new terminals are introduced, except '(' and ')', and they are always introduced in a well-balanced fashion.

@@

We will now note a few things about the words $w' \in \words(t)$ for some given $t \in L(G_t)$:

\begin{itemize}
\item All $w'$ have the same basic word. $G'_w$ is constructed in such a way that the only possible difference is one of parentheses. This also implies that two trees that share a word (i.e., that can be ambiguous) must also share a basic word.
\item Some parentheses are required, i.e., some ranges must be present in all $w'$. For example, removing the parentheses in ''$(1 + 2) * 3$'' changes the tree produced, since multiplication has higher precedence than addition.
\item There is a finite set of possible ranges in the range bags of $w'$. Grouping parentheses can only be added if they exactly cover a node in the parse tree, and other parentheses can only be added where $G_w$ allows them.
\item Duplicated grouping parentheses do not matter. For example, $((1 + 2)) * 3$ permits the same syntax trees as $(1 + 2) * 3$. If there are no parentheses present in $G_w$ then all parentheses are grouping parentheses, thus we can consider the range bag as a \emph{set} instead.
\end{itemize}

@@

This section describes a decision algorithm for detecting unresolvable ambiguities in a language definition, starting with a version with several limitations, most of which are later lifted. The overarching goal is to find a tree $t \in L(G_t)$ such that there is no $w' \in L(G'_w)$ for which $\parse(w') = \{t\}$. Alternatively, find a tree that has no unambiguous words.

@@

\begin{enumerate}
  \item The input language definition contains no parentheses. This implies that all parentheses in $G'_w$ are grouping parentheses, thus we only need to consider words $w' \in L(G'_w)$ whose range bag is a set.

  \item No non-terminals in the input language definition are marked, i.e., there are no forbidden children. This implies that there are no required grouping parentheses.

  \item No production has a right-hand side that matches a single non-terminal, i.e., $\forall (N->i:r) \in D.\ \NT \cap L(r) = \emptyset$\footnote{I haven't really written a proper description for the language definition, but I'm here using $D$ for it, and considering it essentially as a set of ''productions''.}.
\end{enumerate}

\noindent The first two limitations allow us to place all words in $\words(t)$ for some given $t$ in a lattice, whose structure is formed by the subset ordering of the range sets. For example, the two trees in Figure~\ref{fig:ambig-example-square} have the lattices of words seen in Figures~\ref{fig:lattice1} and \ref{fig:lattice2} respectively.

The bottom word corresponds to the set of required parentheses (i.e., the empty set, in this sub-problem), while the top word corresponds to the set of possible parentheses ranges. All trees that can be ambiguous must thus share the same bottom word, while the top may differ.

@@

\begin{itemize}
\item The top word is unambiguous, then this is not a tree we are looking for (since it has at least one unambiguous word).
\item It is ambiguous. That means that there is some other lattice for some other tree that also contains the word. Call the top word of this other lattice $w'_2$. For $w'$ to be in the lattice whose top is $w'_2$ its rangeset must be a subset of the rangeset of the latter. But this is true also for all the other words in the lattice, thus there are no unambiguous words in the lattice, i.e., this tree is what we are looking for.
\end{itemize}

\noindent The algorithm thus looks for two trees, where the set of possible parentheses for one is a superset of the other. To do this, we use a linear representation of each lattice: namely the top word. If two trees have the same top word, that implies that their sets of possible parentheses are equal. If we can add parentheses to one word and get the other word, that implies that the rangeset of the former is a subset of the latter.

We will now construct a pushdown automaton that recognizes these top words in such a way that there is a bijection between successful runs and trees in $L(G_t)$. As a running example, we will use the following (very simple) language definition $D$:

\begin{center}
\begin{tabular}{@{}l@{\quad$->$\quad}l@{ $:$\quad}l@{}}
  \synt{N} & Succ & 's' \synt{N}\\
  \synt{N} & Zero & 'z' \\
\end{tabular}
\end{center}

@@

\noindent Intuitively, we parse a child node, but put parentheses around it, then return. This is where we use assumption 3, without it we might introduce double parentheses here.

@@

We now have a pushdown automaton (call it $A_{()}$) that recognizes the top word for each tree in $L(G_t)$. Next we need to be able to add arbitrary parentheses, to produce a word with a rangeset superset. To do this we create a copy of $A_{()}$ with one modification: for every state $s$ in $A_{()}$ (except the initial and final states), add two transitions $s \xrightarrow{'(', +p} s$ and $s \xrightarrow{')', -p} s$. To avoid cluttering the graph too much, these transitions are shown unlabeled below:

@@

\noindent We will call this automaton $A'_{()}$. Successful runs in this automaton have a surjection to runs in $A_($ (ignore transitions along the newly added edges), and thus also have a surjection to parse trees in $L(G_t)$. We can also note that every sucessful run in $A_{()}$ is also a successful run in $A'_{()}$, since the latter has all states and transitions of the former. Furthermore, two distinct runs, one in $A_{()}$ (call it $p$) and one in $A'_{()}$ (call it $p'$), that both recognize the same word must represent different parse trees in $L(G_t)$. To see why, we consider two cases:

@@

\noindent Two distinct successful runs that accept the same word thus represent two trees where one permits a superset rangeset of the other. To find such runs we construct a product automaton and trim it. We can construct a product automaton since both $A_{()}$ and $A'_{()}$ are visibly pushdown, with the same partition of the input alphabet (push on open parenthesis, pop on close parenthesis, do nothing otherwise). We can trim the product since it retains the same partitioning and thus is also visibly pushdown.

@@

% TODO: this section should be moved to be in the appropriate position {
\newpage

\begin{figure*}[ht]
  \begin{tabular}{@{}l@{\quad$->$\quad}l@{ $:$\quad}l@{}}
    \synt{Expr} & Sum & \synt{Expr} \verb|'+'| \synt{Expr} \\
    \synt{Expr} & Product & \synt{Expr}$_{Sum}$ \verb|'*'| \synt{Expr}$_{Sum}$ \\
    \synt{Expr} & Number & \synt{Number} \\
  \end{tabular}
  \caption{The input language definition used as a running example, a simple expression language with addition and multiplication, with precedence defined, but not associativity. Assumes that \synt{Number} matches a numeric terminal.}
\end{figure*}

\begin{figure*}[ht]
  \begin{tikzcd}
    L(G_w) & L(G_t) \arrow[l, "\yield"] \\
    L(G'_w) & L(G'_t) \arrow[u, "\semantic"'] \arrow[l, "\yield"]
  \end{tikzcd}
  \caption{The generated grammars and their relation to each other.}
\end{figure*}

\begin{figure*}[ht]
  \begin{subfigure}[t]{.45\linewidth}
    \centering
    \begin{tabular}{@{}l@{\quad$->$\quad}l@{}}
      \toprule
      \synt{Expr} & \synt{Expr} \verb|'+'| \synt{Expr} \\
      \synt{Expr} & \synt{Expr} \verb|'*'| \synt{Expr} \\
      \synt{Expr} & \synt{Number} \\
      \bottomrule
    \end{tabular}
    \caption{$G_w$, the generated abstract syntax.}
  \end{subfigure}%
%
  \begin{subfigure}[t]{.54\linewidth}
    \centering
    \begin{tabular}{@{}l@{\quad$->$\quad}lll@{}}
      \toprule
      \synt{Expr} & $[_{Sum}$ & \synt{Expr} \verb|'+'| \synt{Expr} & $]_{Sum}$ \\
      \synt{Expr} & $[_{Product}$ & \synt{Expr} \verb|'*'| \synt{Expr} & $]_{Product}$ \\
      \synt{Expr} & $[_{Number}$ & \synt{Number} & $]_{Number}$ \\
      \bottomrule
    \end{tabular}
    \caption{$G_t$, a language describing a linear encoding of the parse trees of $G_w$.}
  \end{subfigure}

  \begin{subfigure}{.45\linewidth}
    \centering
    \begin{tabular}{@{}l@{\quad$->$\quad}l@{}}
      \toprule
      \synt{Expr} & \synt{Expr} \verb|'+'| \synt{Expr} \\
      \synt{Expr} & \synt{Expr}$_{Sum}$ \verb|'*'| \synt{Expr}$_{Sum}$ \\
      \synt{Expr} & \synt{Number} \\
      \synt{Expr} & \verb|'('| \synt{Expr} \verb|')'| \\
      \midrule
      \synt{Expr}$_{Sum}$ & \synt{Expr}$_{Sum}$ \verb|'*'| \synt{Expr}$_{Sum}$ \\
      \synt{Expr}$_{Sum}$ & \synt{Number} \\
      \synt{Expr}$_{Sum}$ & \verb|'('| \synt{Expr} \verb|')'| \\
      \bottomrule
    \end{tabular}
    \caption{$G'_w$, the generated concrete syntax.}
  \end{subfigure}%
%
  \begin{subfigure}{.54\linewidth}
    \centering
    \begin{tabular}{@{}l@{\quad$->$\quad}lll@{}}
      \toprule
      \synt{Expr} & $[_{Sum}$ & \synt{Expr} \verb|'+'| \synt{Expr} & $]_{Sum}$ \\
      \synt{Expr} & $[_{Product}$ & \synt{Expr}$_{Sum}$ \verb|'*'| \synt{Expr}$_{Sum}$ & $]_{Product}$ \\
      \synt{Expr} & $[_{Number}$ & \synt{Number} & $]_{Number}$ \\
      \synt{Expr} & $[_{Grouping}$ & \verb|'('| \synt{Expr} \verb|')'| & $]_{Grouping}$ \\
      \midrule
      \synt{Expr}$_{Sum}$ & $[_{Sum2}$ & \synt{Expr}$_{Sum}$ \verb|'*'| \synt{Expr}$_{Sum}$ & $]_{Sum2}$ \\
      \synt{Expr}$_{Sum}$ & $[_{Number2}$ & \synt{Number} & $]_{Number2}$ \\
      \synt{Expr}$_{Sum}$ & $[_{Grouping2}$ & \verb|'('| \synt{Expr} \verb|')'| & $]_{Grouping2}$ \\
      \bottomrule
    \end{tabular}
    \caption{$G'_t$, a language describing a linear encoding of the parse trees of $G'_w$.}
  \end{subfigure}%
  \caption{The generated grammars.}
\end{figure*}

\begin{figure*}[ht]
  \begin{subfigure}[b]{.45\linewidth}
    \[1 + 2 * 3\]
    \caption{Example word in $L(G_w)$.}
  \end{subfigure}
  \begin{subfigure}[b]{.45\linewidth}
    \begin{center}
    \Tree [.Product
        [.Sum
          [.Number 1 ]
          '+'
          [.Number 2 ] ]
        '*'
        [.Number 3 ] ]
    \end{center}
    \[[_P [_S [_N 1 ]_N + [_N 2 ]_N ]_S * [_N 3 ]_N ]_P\]
    \caption{Example word in $L(G_T)$ and visualized tree.}
  \end{subfigure}

  \begin{subfigure}[b]{.45\linewidth}
    \[(1 + 2) * 3\]
    \caption{Example word in $L(G'_w)$.}
  \end{subfigure}
  \begin{subfigure}[b]{.45\linewidth}
    \begin{center}
    \Tree [.Product
        [.Grouping2
          '('
          [.Sum
            [.Number 1 ]
            '+'
            [.Number 2 ] ]
          ')' ]
        '*'
        [.Number2 3 ] ]
    \end{center}
    \[[_P [_{G2} ( [_S [_N 1 ]_N + [_N 2 ]_N ]_S ) ]_{G2} * [_{N2} 3 ]_{N2} ]_P\]
    \caption{Example word in $L(G'_T)$ and visualized tree.}
  \end{subfigure}
  \caption{Example with words from each grammar that correspond to each other. The subscripts are abbreviated as follows: $S = \text{Sum}$ , $P = \text{Product}$ , $N = \text{Number}$ , and $G = \text{Grouping}$ .}
\end{figure*}

\begin{figure*}[ht]
  \begin{subfigure}[b]{.45\linewidth}
    \begin{align*}
      1 + 2 + 3
    \end{align*}
    \caption{Example word in $L(G_w)$.}
  \end{subfigure}
  \begin{subfigure}[b]{.45\linewidth}
      \begin{align*}
        [_S           [_S  &[_N 1 ]_N + \hphantom{[_S} [_N 2 ]_N           ]_S  + [_N 3 ]_N \hphantom{]_S} ]_S \\
        [_S \hphantom{[_S} &[_N 1 ]_N +           [_S  [_N 2 ]_N \hphantom{]_S} + [_N 3 ]_N           ]_S  ]_S
      \end{align*}
    \caption{Example words in $L(G_T)$.}
  \end{subfigure}

  \begin{subfigure}[b]{.45\linewidth}
    \begin{align*}
      1 + 2 + 3
    \end{align*}
    \caption{Example word in $L(G'_w)$.}
  \end{subfigure}
  \begin{subfigure}[b]{.45\linewidth}
      \begin{align*}
        [_S            [_S &[_N 1 ]_N + \hphantom{[_S} [_N 2 ]_N            ]_S + [_N 3 ]_N \hphantom{]_S}]_S \\
        [_S \hphantom{[_S} &[_N 1 ]_N +            [_S [_N 2 ]_N \hphantom{]_S} + [_N 3 ]_N ]_S ]_S
      \end{align*}
    \caption{Example words in $L(G'_T)$.}
  \end{subfigure}
  \caption{Example with an ambiguous word in $L(G'_w)$ with corresponding words from the other grammars.}
  \label{fig:ambig-example-square}
\end{figure*}

\begin{figure*}[ht]
  \begin{tikzpicture}[every node/.style={rectangle,draw}]
    \node (a) [draw=white] at (0,5) {$((1) + (2)) + (3)$};

    \node (b1) at (-3.6,4.2) {$(1) + (2) + (3)$};
    \node (b2) [draw=white] at (-1.2,4.2) {$((1) + (2)) + 3$};
    \node (b3) [draw=white] at ( 1.2,4.2) {$((1) + 2) + (3)$};
    \node (b4) [draw=white] at ( 3.6,4.2) {$(1 + (2)) + (3)$};

    \node (c1) at (-5.5,3) {$(1) + (2) + 3$};
    \node (c2) at (-3.3,3) {$(1) + 2 + (3)$};
    \node (c3) [draw=white] at (-1.1,3) {$((1) + 2) + 3$};
    \node (c4) at (1.1,3) {$1 + (2) + (3)$};
    \node (c5) [draw=white] at (3.3,3) {$(1 + (2)) + 3$};
    \node (c6) [draw=white] at (5.5,3) {$(1 + 2) + (3)$};

    \node (d1) at (-3.6,1.8) {$(1) + 2 + 3$};
    \node (d2) at (-1.2,1.8) {$1 + (2) + 3$};
    \node (d3) at ( 1.2,1.8) {$1 + 2 + (3)$};
    \node (d4) [draw=white] at ( 3.6,1.8) {$(1 + 2) + 3$};

    \node (e) at (0, 1) {$1 + 2 + 3$};

  \draw (d1.90) -- (c1.270);
  \draw (d1.90) -- (c2.270);
  \draw (d1.90) -- (c3.270);
  \draw (d2.90) -- (c1.270);
  \draw (d2.90) -- (c4.270);
  \draw (d2.90) -- (c5.270);
  \draw (d3.90) -- (c2.270);
  \draw (d3.90) -- (c4.270);
  \draw (d3.90) -- (c6.270);
  \draw (d4.90) -- (c3.270);
  \draw (d4.90) -- (c5.270);
  \draw (d4.90) -- (c6.270);

  \draw (c1.90) -- (b1.270);
  \draw (c1.90) -- (b2.270);
  \draw (c2.90) -- (b1.270);
  \draw (c2.90) -- (b3.270);
  \draw (c3.90) -- (b2.270);
  \draw (c3.90) -- (b3.270);
  \draw (c4.90) -- (b1.270);
  \draw (c4.90) -- (b4.270);
  \draw (c5.90) -- (b2.270);
  \draw (c5.90) -- (b4.270);
  \draw (c6.90) -- (b3.270);
  \draw (c6.90) -- (b4.270);

  \draw (b1) -- (a);
  \draw (b2) -- (a);
  \draw (b3) -- (a);
  \draw (b4) -- (a);
    %%   \draw[preaction={draw=white, -,line width=6pt}] (a) -- (e) -- (c);
  \end{tikzpicture}
  \caption{The lattice of words for the tree $[_S [_S [_N 1 ]_N + [_N 2 ]_N ]_S + [_N 3 ]_N ]_S$ (in $L(G_t)$). The boxed words are shared with Figure~\ref{fig:lattice2}.}
  \label{fig:lattice1}
\end{figure*}

\begin{figure*}[t]

\begin{tikzpicture}[every node/.style={rectangle,draw}]
  \node (a) [draw=white] at (0,5) {$(1) + ((2) + (3))$};

  \node (b1) at (-3.6,4.2) {$(1) + (2) + (3)$};
  \node (b2) [draw=white] at (-1.2,4.2) {$(1) + ((2) + 3)$};
  \node (b3) [draw=white] at ( 1.2,4.2) {$(1) + (2 + (3))$};
  \node (b4) [draw=white] at ( 3.6,4.2) {$1 + ((2) + (3))$};

  \node (c1) at (-5.5,3) {$(1) + (2) + 3$};
  \node (c2) at (-3.3,3) {$(1) + 2 + (3)$};
  \node (c3) [draw=white] at (-1.1,3) {$(1) + (2 + 3)$};
  \node (c4) at (1.1,3) {$1 + (2) + (3)$};
  \node (c5) [draw=white] at (3.3,3) {$1 + ((2) + 3)$};
  \node (c6) [draw=white] at (5.5,3) {$1 + (2 + (3))$};

  \node (d1) at (-3.6,1.8) {$(1) + 2 + 3$};
  \node (d2) at (-1.2,1.8) {$1 + (2) + 3$};
  \node (d3) at ( 1.2,1.8) {$1 + 2 + (3)$};
  \node (d4) [draw=white] at ( 3.6,1.8) {$1 + (2 + 3)$};

  \node (e) at (0, 1) {$1 + 2 + 3$};

  \draw (d1.90) -- (c1.270);
  \draw (d1.90) -- (c2.270);
  \draw (d1.90) -- (c3.270);
  \draw (d2.90) -- (c1.270);
  \draw (d2.90) -- (c4.270);
  \draw (d2.90) -- (c5.270);
  \draw (d3.90) -- (c2.270);
  \draw (d3.90) -- (c4.270);
  \draw (d3.90) -- (c6.270);
  \draw (d4.90) -- (c3.270);
  \draw (d4.90) -- (c5.270);
  \draw (d4.90) -- (c6.270);

  \draw (c1.90) -- (b1.270);
  \draw (c1.90) -- (b2.270);
  \draw (c2.90) -- (b1.270);
  \draw (c2.90) -- (b3.270);
  \draw (c3.90) -- (b2.270);
  \draw (c3.90) -- (b3.270);
  \draw (c4.90) -- (b1.270);
  \draw (c4.90) -- (b4.270);
  \draw (c5.90) -- (b2.270);
  \draw (c5.90) -- (b4.270);
  \draw (c6.90) -- (b3.270);
  \draw (c6.90) -- (b4.270);

  \draw (b1) -- (a);
  \draw (b2) -- (a);
  \draw (b3) -- (a);
  \draw (b4) -- (a);
%%   \draw[preaction={draw=white, -,line width=6pt}] (a) -- (e) -- (c);
\end{tikzpicture}
\caption{The lattice of words for the tree $[_S [_N 1 ]_N + [_S [_N 2 ]_N + [_N 3 ]_N ]_S ]_S$ (in $L(G_t)$). The boxed words are shared with Figure~\ref{fig:lattice1}.}
\label{fig:lattice2}
\end{figure*}

\clearpage

% }
